Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-11-05 Latex_files/Thesis/chapters/chapter 2/main.tex
praveen commments plus chp 4 rework

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a 
conservative force brought about by the polarisation of dielectric 
materials which is directed towards the point of maximum intensity 
(for a simple Gaussian beam this would be at it's focal point) 
\cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
higher numerical aperture lenses in order to increase the intensity, 
thus allowing for stronger gradient forces at the trade off of 
decreased trapping depth.

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far exceeds both the the scattering force, 
and the thermal energy of the particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient force
will try to recentre the sphere if it is displaced. If the restoring 
force is proportional to the sphere's displacement this is referred 
to as a harmonic trap. The ratio of the restoring force to the 
displacement is called the trap stiffness (denoted by $\kappa$), and 
is an important characteristic of the trap. Shown below is a typical 
force displacement curve for a single sphere in a focused Gaussian beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

@@

\noindent 
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the latter acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

@@

The Fickian model describes the net flux of solute molecules into a 
finite volume of fluid being proportional to the density gradient $\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
solute molecules do not collide with one another or with molecules in 
the solution; overall, the Fickian model is used to describe how solute 
molecules disperse over long periods of time. It does not provide any 
incite into the forces acting on individual molecules, whereas the 
Langevin model captures the physical interactions between an individual 
molecule and the surrounding fluid over a wide range of time scales.

The Einstein model of diffusion expands upon the Fickian model by 
considering the collisions between individual molecules. If we 
consider a single particle suspended in a solution it will experience 
multiple collisions with other molecules \cite{Gillespie2012a} over a 
given time interval $\delta t$. The Einstein molecule allows us to 
consider how individual molecules behave while suspended in a 
surrounding fluid. Where it begins to falter is when we consider 
bringing $\delta t$ to $0$; in this scenario the molecule experiences 
not many but only individual collisions. The Einstein model does not 
consider the inertia of the molecule in question and so for very short 
time frames the model implies that the molecule's velocity changes 
instantly after each collision \cite{Gillespie2012a,Gillespie2012b}. 
Furthermore, the kinetic energy of each collision is not limited by 
the thermal energy of the system, meaning that using the Einstein model 
to predict a particle's trajectory to an finite degree of accuracy 
implies that the particle has infinite kinetic energy \cite{Gillespie2012b}. 
This failure to describe motion over smaller time frames is addressed 
in the Langevin model by accounting for the fluid drag of the system, 
where any sudden change in velocity must result in a proportionally 
opposed drag force \cite{Gillespie2012c}.

@@

Like with the analytical expression of the mean squared displacement we 
see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value 
that when converted to length units represents the maximum displacement 
the particle can move beyond the focus. When $f\gg f_c$ the PSD falls 
off exponentially which denotes the particle is freely diffusing within 
the beam focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{PSD.png}
	\caption{Example PSD fitted using \eqref{eq:lorentzian}, power spectra is collected from an optically trapped silica sphere suspended in water. The difference in magnitude is due to the asymmetry of the quadrant photo diode having a stronger signal response in the direction of the polarisation vector. Using a correction factor (see \eqref{eq:correction_factor}) will adjust the power spectra to better describe the trap shape.}
\end{figure}

@@

\subsubsection{Power spectral analysis for rotating objects}
The scattering from rotating objects can be partially characterised via
the power spectrum density method. Typically papers reporting on the 
rotation rate of a trapped object will collect a power spectrum and look
at the maximum frequency term to determine its rotation rate. However this
often neglects any information on the trapping forces acting on particle
and only looks at the torque applying around one of the particle's primary 
axis. There has only been one notable effort to fully characterise the 
full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
a theoretical model for that relates the power spectrum to both the 
rotational and translational motion simultaneously.

@@

Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
motion is generalised to a 3 dimensional case. Except now, the dimer 
is undergoing random rotational motion in addition to its Brownian 
translational motion the first term on the right hand side is no longer 
purely a function of the dimer's position but also on its orientation. 
The rotational form of the Langevin equation for a dipole within an external 
potential is given as:
\begin{align}
  \frac{{d}\vec{u}(t)}{{dt}}
  =
  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}

@@

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this project we use the work of Vigilante to perform a 
systematic study of the dynamics demonstrated by asymmetric dimers 
in both plane and circularly polarised light. The simulations have 
two primary frames of reference that require explanation. Firstly, 
there is the laboratory frame which describes the coordinate system, 
we define the origin of our simulation as the focus of our trapping 
beam. While the choice in direction for the $x$ and $y$ axis are 
arbitrary the $z$ axis is specifically chosen so that it is aligned 
with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

@@

In order to simulate a typical experimental set up with a QPD 
installed as a position detection system we need to evaluate 
the total magnitude of the electric field incident on the 
photo-diode surface. While trapping a micro-particle, the 
scattered and incident fields combine together and interfere 
with one another. These fields are collected by a condenser 
lens in the far field limit and are focused onto the QPD 
surface, the total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
\end{align}

@@

The incident beam is simple enough to define given our set up 
parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
\textit{Ott} uses a point matching approach to approximate the beam shape 
coefficients of the incident field by fitting it to the far field estimate. 
From the QPD's perspective it is receiving light both from the incident 
and scattered beam simultaneously, as such both fields must be expressed
using outgoing vector spherical harmonics.
\begin{align}
	E_{\rm inc}(r)&=
	E_0\sum^\infty_{nm}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
	\\
	E_{\rm scat}(r)&=E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
	-i p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
	\right] 
\end{align}

We can, if we want, further restrict the incident beam by setting 
the truncation angle to match our microscope object, this applies a 
hard cut off to the incident beam to better represent experimental 
conditions. In order to compute the scattering from 
the target particle \textit{ott} uses the $T$-matrix method, this is 
not essential for a simple sphere but is essential for complex shaped 
particles such as dimers. The matrix from mstm needs to be repackaged 
to work with the \textit{ott} software, converting it from a column 
vector into a system of sub matrices.

The scattered and incident fields are then combined together 
in the far field to get $I(x,y)$, the quadrant and overall signals are 
calculated via:
\begin{align}
	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
\end{align}

@@

\newpage
While translational motion have no knock on effects to the QPD signal, 
rotational changes are between the incident and total fields. This results
in rotational motion being biased in the QPD signal - even when collecting
signals from isotropic scatterers. To remedy this an inverse rotation is
applied to the total field.

@@

The discrepancy between our simulated QPD and the results from 
Rohrbach can be attributed to the fact that the position to 
signal error grows as you move further from trap focus. In most
optical trapping experiments we usually are calibrating a strong
trap where the mean square displacement falls well within the
linear regime shown in fig.~\ref{fig:Rohrbach}. In which case the
error between Rohrbach's model and our own is inconsequential for 
calibration purposes. With this any trajectory can be collected 
from the QPD by displacing and rotating the beam accordingly, 
later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of accuracy
that can be achieved using back focal plane interferometry to 
characterise non-spherical targets.

>>>>

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a 
conservative force brought about by the polarisation of dielectric 
materials which is directed towards the point of maximum intensity 
(for a simple Gaussian beam this would be at it's focal point) 
\cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise a
higher numerical aperture lens in order to increase the intensity, 
thus allowing for stronger gradient forces at the trade off of 
decreased trapping depth.

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far cancels out the the scattering force, 
and the thermal energy of the particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient force
will try to recentre the sphere if it is displaced. If the restoring 
force is proportional to the sphere's displacement this is referred 
to as a harmonic trap. The ratio of the restoring force to the 
displacement is called the trap stiffness (denoted by $\kappa$), and 
is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap}
demonstrates an example case of the force-displacement curve for a 
simple sphere in an optical trap. Laser power and numerical aperture
are omitted as the exact shape is not of importance, only the linear 
region about the origin.
\begin{figure}[h!]
	\centering
	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

@@

\noindent 
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the former acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

@@

The Fickian model describes the net flux of solute molecules into a 
finite volume of fluid being proportional to the density gradient 
$\delta\rho(u,t)/\delta u$ \cite{Gillespie2012}. It assumes that the 
solute molecules do not collide with one another or with molecules in 
the solution; overall, the Fickian model is used to describe how solute 
molecules disperse over long periods of time. It does not provide any 
insight into the forces acting on individual molecules, whereas the 
Langevin model captures the physical interactions between an individual 
molecule and the surrounding fluid over a wide range of time scales.

The Einstein model of diffusion expands upon the Fickian model by 
considering the collisions between individual molecules. If we 
consider a single particle suspended in a solution it will experience 
multiple collisions with other molecules \cite{Gillespie2012a} over a 
given time interval $\delta t$. The Einstein model allows us to 
consider how individual molecules behave while suspended in a 
surrounding fluid. Where it begins to falter is when we consider 
bringing $\delta t$ to $0$; in this scenario the molecule experiences 
not many but only individual collisions. The Einstein model does not 
consider the inertia of the molecule in question and so for very short 
time frames the model implies that the molecule's velocity changes 
instantly after each collision \cite{Gillespie2012a,Gillespie2012b}. 
Furthermore, the kinetic energy of each collision is not limited by 
the thermal energy of the system, meaning that using the Einstein model 
to predict a particle's trajectory to an finite degree of accuracy 
implies that the particle has infinite kinetic energy \cite{Gillespie2012b}. 
This failure to describe motion over smaller time frames is addressed 
in the Langevin model by accounting for the fluid drag of the system, 
where any sudden change in velocity must result in a proportionally 
opposed drag force \cite{Gillespie2012c}.

@@

Like with the analytical expression of the mean squared displacement we 
see two distinct regions, when $f\ll f_c$ the PSD reaches a plateau value 
that when converted to length units represents the maximum displacement 
the particle can move beyond the focus. When $f\gg f_c$ the PSD falls 
off exponentially which denotes the particle is freely diffusing within 
the beam focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{PSD.png}
	\caption{Example PSD fitted using \eqref{eq:lorentzian}, power 
		spectra is collected from an optically trapped silica sphere 
		suspended in water. The difference in magnitude is due to 
		the asymmetry of the quadrant photo diode having a stronger 
		signal response in the direction of the polarisation vector. 
		Using a conversion factor (see \eqref{eq:conversion_factor}) 
		will adjust the power spectra to better describe the trap shape.}
\end{figure}

@@

\subsubsection{Power spectral analysis for rotating objects}
The scattering from rotating objects can be partially characterised via
the power spectrum density method. Typically papers reporting on the 
rotation rate of a trapped object will collect a power spectrum and look
at the maximum frequency term to determine its rotation rate. However this
often neglects any information on the trapping forces acting on particle
and only looks at the torque applying around one of the particle's primary 
axis. There has only been one notable effort to fully characterise the 
full trapping dynamics on a rotating body, \cite{Yogesha2012} developed 
a theoretical model that relates the power spectrum to both the 
rotational and translational motion simultaneously.

@@

Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational 
motion is generalised to a 3 dimensional case. Here we omit the inertia term 
as the restoration time of the optical trap is far greater than the characteristic
time of the particle ($\kappa/\gamma \gg m/\gamma$). Now, the dimer 
is undergoing random rotational motion in addition to its Brownian 
translational motion the first term on the right hand side is no longer 
purely a function of the dimer's position but also of its orientation. 
The rotational form of the Langevin equation for a dipole within an external 
potential is given as:
\begin{align}
  \frac{{d}\vec{u}(t)}{{dt}}
  =
  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}

@@

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this project we use the work of Vigilante \cite{Vigilante2020} 
to perform a systematic study of the dynamics demonstrated by 
asymmetric dimers in both plane and circularly polarised light. 
The simulations have two primary frames of reference that require 
explanation. Firstly, there is the laboratory frame which 
describes the coordinate system, we define the origin of our 
simulation as the focus of our trapping beam. While the choice in 
direction for the $x$ and $y$ axis are arbitrary the $z$ axis is 
specifically chosen so that it is aligned with the direction of 
beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and 
		particle frame for a dimer and a Gaussian beam. The origin 
		of the laboratory frame is defined as the focus of the 
		trapping beam, with the positive $z$-axis being aligned with 
		the direction of beam propagation. The particle frame is 
		centred on the dimer's centre of mass with the primary axis 
		$U_z$ being aligned with the centres of the two spheres. Black 
		line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

@@

In order to simulate a typical experimental set up with a QPD 
installed as a position detection system we need to evaluate 
the total magnitude of the electric field incident on the 
photo-diode surface. While trapping a micro-particle, the 
scattered and incident fields combine together and interfere 
with one another. These fields are collected by a condenser 
lens in the far field limit and are focused onto the QPD 
surface, the total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(\theta_{NA}-\theta(x,y))
\end{align}

@@

The incident beam is simple enough to define given our set up 
parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
\textit{Ott} uses a point matching approach to approximate the beam shape 
coefficients of the incident field by fitting it to the far field estimate. 
From the QPD's perspective it is receiving light both from the incident 
and scattered beam simultaneously, as such both fields must be expressed
using outgoing vector spherical harmonics.
\begin{align}
	E_{\rm inc}(r)&=
	E_0\sum^\infty_{n}\sum^n_{m=-n}\left[a_{mn}{\bf M}^{(2)}_{nm}({\bf r})+
	b_{nm}{\bf N}^{(2)}_{nm}({\bf r}) \right] 
	\\
	E_{\rm scat}(r)&=E_0 \sum^\infty_n \sum^n_{m=-n} \left[
	p_{nm}{\bf N}_{nm}^{(2)}({\bf r})+q_{nm}{\bf M}_{nm}^{(2)}({\bf r})
	\right] 
\end{align}

Where the superscript (2) denotes an outgoing spherical harmonic
function. In order to compute the scattering from the target particle \textit{ott} uses the $T$-matrix method, this is not essential 
for a simple sphere but is essential for complex shaped particles 
such as dimers. The scattered and incident fields are then combined 
together in the far field to get $I(x,y)$, the quadrant and overall 
signals are calculated via:
\begin{align}
	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
\end{align}

@@

While translational motion have no knock on effects to the QPD signal, 
rotational changes are between the incident and total fields. This results
in rotational motion being biased in the QPD signal - even when collecting
signals from isotropic scatterers. To remedy this an inverse rotation is
applied to the total field.

@@

The discrepancy between our simulated QPD and the results 
from Rohrbach can be attributed to the fact that the position 
to signal error grows as you move further from trap focus. 
As the particle moves further from the trap focus the error 
in the attributed signal grows substantially \cite{Rohrbach2002}.
In most optical trapping experiments we usually are calibrating 
a strong trap where the mean square displacement falls well within 
the linear regime shown in fig.~\ref{fig:Rohrbach}. In which case 
the error between Rohrbach's model and our own is inconsequential 
for calibration purposes. With this any trajectory can be collected 
from the QPD by displacing and rotating the beam accordingly, 
later in chapters \ref{chapter:langevin_dynamics} and \ref{chapter:simulated_detection} we discuss the limits of 
accuracy that can be achieved using back focal plane interferometry 
to characterise non-spherical targets.
